\section{Overview}

This report is divided into two principal sections: 
\begin{enumerate}
    \item \textbf{Research}: Discusses both current and projected research activities as part of this PhD program.
    \item \textbf{PhD Study Programme}: Details progress and future objectives aligned with the study requirements of the programme.
\end{enumerate}

\subsection*{Aim}

This PhD aims to enhance the understanding and transparency of machine learning 
predictions through the introduction of novel methods and evaluations of explainable AI (XAI).

\subsection*{Current or planned contributions}
The contributions are categorised as follows: Papers that have been submitted are highlighted 
in \textbf{bold}, and those that are ongoing or planned are in \textit{italics}:
\begin{itemize}
    \item \textbf{Novel Explanation Methods}:
    \begin{itemize}
        \item \textbf{XAIStories}: Natural language narratives formed from SHAP and counterfactual explanations.
        \item \textbf{CoF Tables}: A method to aggregate instance-level explanations to identify shortcuts in 
        image classification models.
        \item \textit{CoF Tables 2.0}: Improving CoF tables through combinations of segments, 
        new segmentation and edit functions. Ablation studies to measure effectiveness of different components.
    \end{itemize}
    \item \textbf{Evaluation of XAI Methods}:
    \begin{itemize}
        \item \textit{PCherry}: Evaluates the probablilty that a given set of counterfactuals includes 
        cherry-picked explanations.
        \item \textit{PCherry Fairness Spin-Off}: Examines the propensity of counterfactual explanations to 
        fairwash through optimisations aimed at enhancing plausibility and sparsity.
    \end{itemize}
    \item \textit{AXA Collaboration}: Improving mail switching LLM through the analysis of counterfactual explanations.
    \item \textit{Fairness in Image Classification}: An investigation into the fairness of image classification 
    models using CoF tables.
\end{itemize}

\subsection*{Progress towards Study Program}
The program requires the accumulation of 30 credits. Below is the summary of credits earned to date:

\begin{tabular}{|c|l|c|c|}
    \hline
    \textbf{Competency} & \textbf{Name}             & \textbf{Credits} & \textbf{Date}   \\ \hline
    A                   & DLCV                      & 6                & Jun 23         \\ \hline
    C                   & Masters Supervision       & 1                & Jan 23        \\ \hline
    C                   & Masters Supervision       & 1                & Jan 24        \\ \hline
    D                   & XAIStories arXiv          & 2                & Sep 23        \\ \hline
    \multicolumn{3}{|r|}{\textbf{Total Credits}}   & 10               \\ \hline
\end{tabular}